Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-04-20 apj.tex
finish new draft

\begin{abstract}
\color{orange}
Empirical modeling has led to precise determinations of the peak absolute magnitudes of Type~Ia supernovae (SNe~Ia), making this class
of object an excellent distance indicator.  Improved modeling of SN~Ia absolute magnitudes may account for physical effects that
are expected but not included
in current models, lead to more precise per-supernova distances, and further constrain systematic uncertainties.  In this article, we present the
results of    
\color{black}
an empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
as applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and the color parameters, to better than 
%-----
$(1-5\times10^{-5})$
%------
confidence
\color{orange}
based on the posterior.
\color{black}
The color changes allowed by the model are almost parallel to the color changes allowed by the extinction model of  \citet{1999PASP..111...63F} with floating
total-to-selective extinction parameter $R^F$.  The deduced color excess
is consistent with an effective $R^F = 2.44$.
The derived dust curve is able to reproduce the optical colors of the highly reddened SN~2014J better than the \citet{1999PASP..111...63F} dust model.
Inspired that our  two-parameter model, with no conditioning that would steer it toward a dust model, produces behavior similar to that expected from Galactic dust, we extend  the model to three color parameters.
In this new case we obtain consistent results for
all the parameters in common,
plus a detection of a third parameter at
%--- table.py
$(1-5\times10^{-4})$
%-----
confidence;
this new parameter represents supernova color diversity
beyond that tracked by the spectral features and two-parameter dust models.
%A direct incorporation of dust models into our formalism is impossible. An indirect incorporation of dust
%yields in constant-$R^F$ model $R^F \sim 3$, but the implementation is sensitive to systematic uncertainties that arise
%from the need to specify the colors of an unextincted supernova and its use of absolute extinction, neither of which is used in our original model.
\end{abstract}

@@

\color{orange}
We impose
\begin{equation}
\langle \Delta \rangle=0, \langle k_0 \rangle=0, \langle k_1 \rangle=0.
\end{equation}
These conditions specify the definition of absolute magnitude and zero color relative to which the color excess is measured:
in a linear model these values are arbitrary.
\color{orange}
We impose 
\begin{equation}
\gamma^0_U > 0, \gamma^1_U < 0.
\end{equation}
to exclude degenerate posterior space
associated
with the $\pm$ sign, i.e.\ simultaneous sign flips of
$\gamma^0$--$k_0$ and $\gamma^1$--$k_1$
(when running our fits, $k_0$ and $k_1$ are set to
identical initial conditions, so the difference in the signs of $\gamma^0_U$ and $\gamma^1_U$ distinguishes the two.)

Under exchange of the 0 and 1 indices, $\gamma^0$--$k_0$ and $\gamma^1$--$k_1$ are degenerate with
each other.  That is, the solution for $\gamma^0$ of a subset of the chains is the solution for $\gamma^1$ of the remaining chains, and vice versa.
We find that a putting initial conditions for $\gamma^0$ that are consistent across all chains is sufficient to break the degeneracy of our mixture model.
\color{black}

@@

\color{orange}
Observed versus deduced colors for four selected colors, and the histogram of their differences,
\color{black}
are shown in Figure~\ref{residual:fig}.
There is no evidence of a catastrophic fit nor of extreme
outlying data.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/residual.pdf} 
            \caption{
            \color{orange}
            Observed versus deduced colors for four selected colors.
On the right for each color is plotted the histogram of the difference between the observations and deductions.
\color{black}
            \label{residual:fig}}
\end{figure}

@@

The histogram of the medians of the grey offsets, $\Delta$, for all supernovae is shown in Figure~\ref{hist:fig}.  The distribution is non-Gaussian, 
has a standard deviation of
%-----
$0.10$
%-----
mag, and a broad tail in the positive (fainter) direction. 
\color{orange}
Of the three extreme outliers, the supernovae with largest and third largest $\Delta$'s have  heliocentric redshifts of 
 $0.0015$ and $0.0085$ respectively, and so have peculiar velocities that can contribute significantly to $\Delta$.
\color{black}
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/Delta_hist.pdf} 
   \caption{Normalized ideogram and histogram of the medians of the grey offset $\Delta$. 
   \label{hist:fig}}
\end{figure}

Non-trivial residual magnitude dispersions are captured in $C_c$.  The diagonal elements are captured by the $\sigma$ parameters;
the residual intrinsic dispersion ranges from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.
\color{orange}
Of particular interest in $C_c$ is intrinsic color dispersion.  The intrinsic dispersions in $V$ and colors with respect to $V$ are
\begin{equation}
\sigma(V,U-V,B-V,V-R,V-I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
\left(
0.021^{+0.004}_{-0.005} ,
0.067^{+0.010}_{-0.010},
0.034^{+0.005}_{-0.005},
0.023^{+0.002}_{-0.002},
0.053^{+0.004}_{-0.004}
\right)
%\end{array}
% \end{pmatrix} 
 \text{mag}.
 \label{sig_intrinsic:eqn}
 \end{equation}
The correlations between these parameters are
\begin{equation}
Cor(V,U-V,B-V,V-R,V-I)=
\begin{pmatrix}
\begin{array}{rrrrr}
1 & -0.48^{+0.20}_{-0.18} & -0.34^{+0.21}_{-0.22} & 0.88^{+0.10}_{-0.25} & 0.60^{+0.11}_{-0.19} \\
-0.48^{+0.20}_{-0.18} & 1 & 0.60^{+0.10}_{-0.13} & -0.38^{+0.18}_{-0.17} & -0.33^{+0.15}_{-0.14} \\
-0.34^{+0.21}_{-0.22} & 0.60^{+0.10}_{-0.13} & 1 & -0.21^{+0.17}_{-0.17} & 0.08^{+0.14}_{-0.15} \\
0.88^{+0.10}_{-0.25} & -0.38^{+0.18}_{-0.17} & -0.21^{+0.17}_{-0.17} & 1 & 0.65^{+0.06}_{-0.07} \\
0.60^{+0.11}_{-0.19} & -0.33^{+0.15}_{-0.14} & 0.08^{+0.14}_{-0.15} & 0.65^{+0.06}_{-0.07} & 1 \\
\end{array}
 \end{pmatrix}.
 \label{cor_intrinsic:eqn}
 \end{equation}

\color{black}
%
%Given that the 
%off-diagonal elements of $C_c$ are parameterized by the Cholesky factors of a correlation matrix rather than the matrix elements themselves,
%it is not straightforward to present correlation matrix confidences:
%the average over Monte Carlo links of the
%Cholesky factors will not  generally yield a correlation matrix.  
%To characterize a typical posterior draw of $C_c$ we use the matrix that is the mean of all covariance realizations in the
%chain, element by element.
%For $UBVRI$ the matrix is
%\begin{equation}
%C_c(U,B,V,R,I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
%0.0038 & 0.0010 & -0.0002 & -0.0000 & 0.0003 \\
%0.0010 & 0.0011 & 0.0002 & -0.0000 & -0.0006 \\
%-0.0002 & 0.0002 & 0.0004 & 0.0000 & -0.0002 \\
%-0.0000 & -0.0000 & 0.0000 & 0.0002 & 0.0002 \\
%0.0003 & -0.0006 & -0.0002 & 0.0002 & 0.0020
%\end{array}
% \end{pmatrix} \text{mag}^2.
% \label{mag_cov:eqn}
% \end{equation}
%The average correlation matrix is
%\begin{equation}
%Cor(U,B,V,R,I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
%1.00 & 0.42 & -0.18 & -0.05 & 0.12 \\
%0.42 & 1.00 & 0.26 & -0.09 & -0.39 \\
%-0.18 & 0.26 & 1.00 & 0.09 & -0.22 \\
%-0.05 & -0.09 & 0.09 & 1.00 & 0.27 \\
%0.12 & -0.39 & -0.22 & 0.27 & 1.00
%\end{array}
% \end{pmatrix}.
% \end{equation}
%The  covariance of the colors $U-V$, $B-V$, $V-R$, and $V-I$ is
%expressed as the standard deviations and
%correlation matrix
% \begin{equation}
% \sigma(U-V, B-V, V-R, V-I)=
% \begin{pmatrix}
%0.068 & 0.034 & 0.023 & 0.053
%  \end{pmatrix} \text{mag},
% \label{color_sd:eqn}
% \end{equation}
% \begin{equation}
% Cor(U-V, B-V, V-R, V-I)=
%\begin{pmatrix}
%\begin{array}{rrrr}
%1.000 & 0.602 & -0.353 & -0.317 \\
%0.602 & 1.000 & -0.197 & 0.082 \\
%-0.353 & -0.197 & 1.000 & 0.647 \\
%-0.317 & 0.082 & 0.647 & 1.000
%\end{array}
%\end{pmatrix}.
%  \label{color_cor:eqn}
% \end{equation}
These color variances are smaller than those derived by \citet{2003A&A...404..901N} or \citet{2007ApJ...659..122J}, for cases where direct comparison can be made.
\color{orange}
%model11_rubin.py
Having no intrinsic color dispersion is disfavored, and not allowing for them can have important consequences.  Running an analysis assuming no intrinsic dispersion (replacing the probabilistic intrinsic magnitudes in Eqn.~\ref{ewsiv:eqn}
with a strict equality), we find significant shifts in the inferred slopes and smaller statistical uncertainties.
\color{black}

@@

Our model can be expressed as
\begin{equation}
A_X = \frac{\gamma^0_X}{\gamma^0_B-\gamma^0_V}  E_{\gamma^0}(B-V) +  \frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V}  E_{\gamma^1}(B-V).
\end{equation}
Our median-fit values for the $\gamma$ coefficients,
%---- fitz11.py
$\frac{\gamma^0_X}{\gamma^0_B-\gamma^0_V}  =\{4.80 ,   3.89,   2.89,   2.22,   1.59\}$ and
$ \frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V}=\{-3.83 ,  -4.42,  -5.42,  -5.15,  -4.71\}$,
%----
can be transformed to a basis composed of the $a(X)$ and $b(X)$ vectors plus a residual vector perpendicular to $a$ and $b$.
The result is
\begin{equation}
\begin{pmatrix}
 \frac{\gamma^0_X}{\gamma^0_B-\gamma^0_V} \\
\frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V} 
\end{pmatrix}=
\begin{pmatrix}
\begin{array}{rr}
1.15 & 2.82 \\
0.72 & -5.27
\end{array}
\end{pmatrix} 
\begin{pmatrix}
a(X) \\
b(X)
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_a \\
\vec{\epsilon}_b
\end{pmatrix},
\label{trans_I:eqn}
\end{equation}
where the two residual vectors
\begin{align}
\vec{\epsilon}_a &=\{0.06, -0.07, -0.07,  0.07, 0.02\}, \\
\vec{\epsilon}_b & =\{-0.06, 0.16, -0.22, 0.33, -0.26\}
\label{res_I:eqn}
\end{align}
contribute  2\% and 5\% respectively to the total  length of the $\gamma$ vectors.

@@

\color{orange}
The preferred values for $\kappa^{-1}$ can be determined from the matrix elements of Eqn.~\ref{model11trans:eqn}: $\kappa^{-1}_1=2.82/1.15 =2.44$, $\kappa^{-1}_2=-5.27/0.72=-7.31$.
The directions of the $\gamma$ vectors, and  the expectations from  \citetalias{1999PASP..111...63F} with $R=\kappa^{-1}$, are shown in 
Figure~\ref{plane:fig}.  There is almost perfect alignment between $\gamma^0$ and $R=2.44$ dust.
 From the Monte Carlo chains, we find a 68\% credible interval for $\kappa^{-1}_1$ of 2.26--2.63.

@@

\color{orange}
As a check of the above interpretation, we generate ten simulated datasets with the same numbers and uncertainties as our data, based on a
SALT2 $x_1=0$, $c=0$ template and  \citetalias{1999PASP..111...63F} dust.  The supernovae
are assigned $E^F(B-V)$ parameters equal to the $B-V$ color excesses determined in our sample
(making an approximative correspondence between observed colors measured in our top-hat filters and the 
model parameter $E^F(B-V)$) , and $R^F$ values drawn from a log-normal
distribution with a mean of $\ln{2.5}$ and 0.05 standard deviation.  The $\alpha$, $\beta$, and $\eta$ parameters are set to zero, so that dust and
measurement uncertainty are the only sources of variation in magnitudes.  The input $R^F=2.5$ is centered
within the 68\% credible interval for $\kappa^{-1}_1$ of 2.38--2.69 from the results of the stacked 10 simulations.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/sn2014j.pdf} 
   \caption{Top: Measured color excess for SN~2014J and the best-fit predictions from this article.   Bottom: Residual color excess
   between the data and model, with a dotted line at zero shown for reference.
   This can be compared with the best-fit results of
   \citet{2014ApJ...788L..21A} shown in their Figure~3.
   \label{sn2014j:fig}}
\end{figure}

The best-fit parameters for SN~2014J are 
%----
$e_0= 2.60$, $ e_1=-1.89$ with covariance
%----
\begin{equation}
\begin{pmatrix}
\begin{array}{rr}
0.050 & 0.058 \\
0.058 & 0.315
\end{array}
\end{pmatrix}.
\end{equation}
The predicted values of observed $E(B_o-V_o)$ from the fit are shown in Figure~\ref{sn2014j:fig}, where they are found to
overlap with the data within the error bars.   
Our results can be compared with those of  \citet{2014ApJ...788L..21A} (Figure 3), who 
use
UV through NIR data to fit \citet{1999PASP..111...63F} dust with $R_V^F=1.4$ and $E(B-V)=1.37$.
Our model predictions provide a significantly better match to the data
despite the only approximate correspondence of the synthetic SNfactory and \citet{2014ApJ...788L..21A} observer passbands.

\color{orange}
Our model and its fit parameters can be associated with the extinction model of  \citetalias{1999PASP..111...63F} using Eqn.\ref{trans_I:eqn}
and the assumption that  SN~2011fe has zero extinction.  The inferred  \citetalias{1999PASP..111...63F} parameters
are $A^F_V=1.28$ and $E^F(B-V)=0.77$ (which together imply $R^F=1.66$) with covariance
\begin{equation}
\begin{pmatrix}
\begin{array}{rr}
0.007 & -0.004 \\
-0.004 & 0.430
\end{array}
\end{pmatrix}.
\end{equation}
\color{black}

@@

@@

The resulting credible intervals of our parameters are given in Table~\ref{global2:tab}.  Comparison with the
results of the 2-parameter extrinsic model given in Table~\ref{global:tab} shows that  parameters that are common to both
are consistent well within 1$\sigma$ uncertainties.  The most significant change is in the slight reduction of the standard
deviations of the magnitude residuals $\sigma_X$.
The newly fitted $\gamma$ parameters have an almost identical association with the  \citetalias{1999PASP..111...63F} dust model, in that now
\begin{equation}
\begin{pmatrix}
 \frac{\gamma^0_X}{\gamma^0_1-\gamma^0_2} \\
\frac{\gamma^1_X}{\gamma^1_1-\gamma^1_2} 
\end{pmatrix}=
\begin{pmatrix}
\begin{array}{rr}
1.15 & 2.79 \\
0.71 & -4.95
\end{array}
\end{pmatrix} 
\begin{pmatrix}
a(X) \\
b(X)
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_a \\
\vec{\epsilon}_b
\end{pmatrix},
\end{equation}
where the residual vectors 
\begin{align}
\vec{\epsilon}_a &=\{0.05, -0.06, -0.06,  0.06,  0.07\},\\
\vec{\epsilon}_b &=\{ -0.10, 0.20,  -0.19,  0.32, -0.30\}
\end{align}
\label{model11trans:eqn}
are perpendicular to the  \citetalias{1999PASP..111...63F} plane and
contribute similar 2\% and 4\% to the total  length of the vector.
(These can be compared to the results of Model I in Eqs.\ \ref{trans_I:eqn}--\ref{res_I:eqn}.)
%The effective total-to-selective extinction for the supernova sample has a slight shift to $R^F_V=2.23 \pm 0.16$.

@@

\color{black}
The analysis is validated using simulated datasets with and without third color parameters: the input parameters are recovered within  uncertainties.
We present here the results of one of the validation tests.  The input values of the supernova parameters are the median values from the analysis of real data.
Ten realizations of simulated data are generated based on the data covariances, and then analyzed separately.   The posteriors of the ten realizations
are aggregated and shown in Figures~\ref{simglobal1:fig}--\ref{simglobal5:fig}.  The input parameters are recovered with negligible bias
relative to the statistical uncertainties.   Note that some bias is expected: the priors for the distributions of $\Delta$, $k$, and $D$ are inconsistent
with the extracted distributions.
%There is a tendency for the slope parameters to be slightly underestimated, which is an expected consequence of 
%the uncertainties in the predictor variable, which in turn lead to 
%regression dilution.  Our analysis is susceptible to this bias because we consciously do not include distributions
%of supernova parameters as part of our model.

>>>>

\begin{abstract}
\color{orange}
Through empirical modeling, the peak absolute magnitude of a Type~Ia supernova (SN~Ia) can be accurately determined 
based on its observed signal, making SNe~Ia  excellent distance indicators.  Improved modeling of SN~Ia absolute magnitudes can account for
further physical
diversity that is expected but not included
in current models, and so consequently can lead to more precise per-object distances with tighter systematic uncertainties.  In this article, we present the
results of    
\color{black}
an empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
as applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and the color parameters, to better than 
%-----
$(1-5\times10^{-5})$
%------
confidence
\color{orange}
based on the posterior.
\color{black}
The color changes allowed by the model are almost parallel to the color changes allowed by the extinction model of  \citet{1999PASP..111...63F} with floating
total-to-selective extinction parameter $R^F$.  The deduced color excess
is consistent with an effective $R^F = 2.44$.
The derived dust curve is able to reproduce the optical colors of the highly reddened SN~2014J better than the \citet{1999PASP..111...63F} dust model.
Inspired that our  two-parameter model, with no conditioning that would steer it toward a dust model, produces behavior similar to that expected from Galactic dust, we extend  the model to three color parameters.
In this new case we obtain consistent results for
all the parameters in common,
plus a detection of a third parameter at
%--- table.py
$(1-5\times10^{-4})$
%-----
confidence;
this new parameter represents supernova color diversity
beyond that tracked by the spectral features and two-parameter dust models.
%A direct incorporation of dust models into our formalism is impossible. An indirect incorporation of dust
%yields in constant-$R^F$ model $R^F \sim 3$, but the implementation is sensitive to systematic uncertainties that arise
%from the need to specify the colors of an unextincted supernova and its use of absolute extinction, neither of which is used in our original model.
\end{abstract}

@@

\color{orange}
We impose
\begin{equation}
\langle \Delta \rangle=0, \langle k_0 \rangle=0, \langle k_1 \rangle=0.
\end{equation}
These conditions specify the definition of absolute magnitude and zero color relative to which the color excess is measured:
in our linear model these values are unconstrained.
\color{orange}
We impose 
\begin{equation}
\gamma^0_U > 0, \gamma^1_U < 0.
\end{equation}
to exclude degenerate posterior space
associated
with the $\pm$ sign, i.e.\ simultaneous sign flips of
$\gamma^0$--$k_0$ and $\gamma^1$--$k_1$
(when running our fits, $k_0$ and $k_1$ are set to
identical initial conditions, so the difference in the signs of $\gamma^0_U$ and $\gamma^1_U$ distinguishes the two.)

The combinations $\gamma^0$--$k_0$ and $\gamma^1$--$k_1$ are degenerate with
each other under exchange of the 0 and 1 indices. For a cleaner presentation of our results, 
we break that degeneracy  by setting consistent initial conditions that succeed in keeping indices consistent over all analysis chains.  
\color{black}

@@

\color{orange}
For four selected colors, a plot of the
observed versus deduced colors and the histogram of their differences
\color{black}
are shown in Figure~\ref{residual:fig}. 
There is no evidence of a catastrophic fit nor of extreme
outlying data.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/residual.pdf} 
            \caption{
            \color{orange}
            Observed versus deduced colors for four selected colors.  Plotted on the right for each color
is the histogram of the difference between the observed and deduced colors.
\color{black}
            \label{residual:fig}}
\end{figure}

@@

The histogram of the medians of the grey offsets, $\Delta$, for all supernovae is shown in Figure~\ref{hist:fig}.  The distribution is non-Gaussian, 
has a standard deviation of
%-----
$0.10$
%-----
mag, and a broad tail in the positive (fainter) direction. 
\color{orange}
Of the three extreme outliers, the supernovae with largest and third largest $\Delta$'s have  heliocentric redshifts of 
 $0.0015$ and $0.0085$ respectively, and so have peculiar velocities that can cause a significant magnitude offset.
\color{black}
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/Delta_hist.pdf} 
   \caption{Normalized ideogram and histogram of the medians of the grey offset $\Delta$. 
   \label{hist:fig}}
\end{figure}

Non-trivial residual magnitude dispersions are captured in $C_c$.  The diagonal elements are captured by the $\sigma$ parameters;
the residual intrinsic dispersion ranges from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.
\color{orange}
Of particular interest within $C_c$ is intrinsic color dispersion.  The intrinsic dispersions in $V$ and colors with respect to $V$ are
\begin{equation}
\sigma(V,U-V,B-V,V-R,V-I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
\left(
0.021^{+0.004}_{-0.005} ,
0.067^{+0.010}_{-0.010},
0.034^{+0.005}_{-0.005},
0.023^{+0.002}_{-0.002},
0.053^{+0.004}_{-0.004}
\right)
%\end{array}
% \end{pmatrix} 
 \text{mag}.
 \label{sig_intrinsic:eqn}
 \end{equation}
The correlations between these parameters are
\begin{equation}
Cor(V,U-V,B-V,V-R,V-I)=
\begin{pmatrix}
\begin{array}{rrrrr}
1 & -0.48^{+0.20}_{-0.18} & -0.34^{+0.21}_{-0.22} & 0.88^{+0.10}_{-0.25} & 0.60^{+0.11}_{-0.19} \\
-0.48^{+0.20}_{-0.18} & 1 & 0.60^{+0.10}_{-0.13} & -0.38^{+0.18}_{-0.17} & -0.33^{+0.15}_{-0.14} \\
-0.34^{+0.21}_{-0.22} & 0.60^{+0.10}_{-0.13} & 1 & -0.21^{+0.17}_{-0.17} & 0.08^{+0.14}_{-0.15} \\
0.88^{+0.10}_{-0.25} & -0.38^{+0.18}_{-0.17} & -0.21^{+0.17}_{-0.17} & 1 & 0.65^{+0.06}_{-0.07} \\
0.60^{+0.11}_{-0.19} & -0.33^{+0.15}_{-0.14} & 0.08^{+0.14}_{-0.15} & 0.65^{+0.06}_{-0.07} & 1 \\
\end{array}
 \end{pmatrix}.
 \label{cor_intrinsic:eqn}
 \end{equation}
\color{black}
%
%Given that the 
%off-diagonal elements of $C_c$ are parameterized by the Cholesky factors of a correlation matrix rather than the matrix elements themselves,
%it is not straightforward to present correlation matrix confidences:
%the average over Monte Carlo links of the
%Cholesky factors will not  generally yield a correlation matrix.  
%To characterize a typical posterior draw of $C_c$ we use the matrix that is the mean of all covariance realizations in the
%chain, element by element.
%For $UBVRI$ the matrix is
%\begin{equation}
%C_c(U,B,V,R,I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
%0.0038 & 0.0010 & -0.0002 & -0.0000 & 0.0003 \\
%0.0010 & 0.0011 & 0.0002 & -0.0000 & -0.0006 \\
%-0.0002 & 0.0002 & 0.0004 & 0.0000 & -0.0002 \\
%-0.0000 & -0.0000 & 0.0000 & 0.0002 & 0.0002 \\
%0.0003 & -0.0006 & -0.0002 & 0.0002 & 0.0020
%\end{array}
% \end{pmatrix} \text{mag}^2.
% \label{mag_cov:eqn}
% \end{equation}
%The average correlation matrix is
%\begin{equation}
%Cor(U,B,V,R,I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
%1.00 & 0.42 & -0.18 & -0.05 & 0.12 \\
%0.42 & 1.00 & 0.26 & -0.09 & -0.39 \\
%-0.18 & 0.26 & 1.00 & 0.09 & -0.22 \\
%-0.05 & -0.09 & 0.09 & 1.00 & 0.27 \\
%0.12 & -0.39 & -0.22 & 0.27 & 1.00
%\end{array}
% \end{pmatrix}.
% \end{equation}
%The  covariance of the colors $U-V$, $B-V$, $V-R$, and $V-I$ is
%expressed as the standard deviations and
%correlation matrix
% \begin{equation}
% \sigma(U-V, B-V, V-R, V-I)=
% \begin{pmatrix}
%0.068 & 0.034 & 0.023 & 0.053
%  \end{pmatrix} \text{mag},
% \label{color_sd:eqn}
% \end{equation}
% \begin{equation}
% Cor(U-V, B-V, V-R, V-I)=
%\begin{pmatrix}
%\begin{array}{rrrr}
%1.000 & 0.602 & -0.353 & -0.317 \\
%0.602 & 1.000 & -0.197 & 0.082 \\
%-0.353 & -0.197 & 1.000 & 0.647 \\
%-0.317 & 0.082 & 0.647 & 1.000
%\end{array}
%\end{pmatrix}.
%  \label{color_cor:eqn}
% \end{equation}
These color variances are smaller than those derived by \citet{2003A&A...404..901N} or \citet{2007ApJ...659..122J}, for cases where direct comparison can be made.
\color{orange}
%model11_rubin.py
Having no intrinsic color dispersion is disfavored, and not allowing for them can have important consequences.  Running an analysis assuming no intrinsic dispersion (replacing the probabilistic intrinsic magnitudes in Eqn.~\ref{ewsiv:eqn}
with a strict equality), we find significant shifts in the inferred slopes and smaller statistical uncertainties.
\color{black}

@@

Our model can be expressed as
\begin{equation}
A_X = \frac{\gamma^0_X}{\gamma^0_B-\gamma^0_V}  E_{\gamma^0}(B-V) +  \frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V}  E_{\gamma^1}(B-V).
\end{equation}
Our median-fit values for the $\gamma$ coefficients,
%---- fitz11.py
$\frac{\gamma^0_X}{\gamma^0_B-\gamma^0_V}  =\{4.80 ,   3.89,   2.89,   2.22,   1.59\}$ and
$ \frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V}=\{-3.83 ,  -4.42,  -5.42,  -5.15,  -4.71\}$,
%----
can be transformed to a basis composed of the $a(X)$ and $b(X)$ vectors plus a residual vector perpendicular to $a$ and $b$.
The result is
\begin{equation}
\begin{pmatrix}
 \frac{\gamma^0_X}{\gamma^0_B-\gamma^0_V} \\
\frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V} 
\end{pmatrix}=
\begin{pmatrix}
\begin{array}{rr}
2.82 & 1.15  \\
-5.27 & 0.72
\end{array}
\end{pmatrix} 
\begin{pmatrix}
a(X) \\
b(X)
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_a \\
\vec{\epsilon}_b
\end{pmatrix},
\label{trans_I:eqn}
\end{equation}
where the two residual vectors
\begin{align}
\vec{\epsilon}_a &=\{0.06, -0.07, -0.07,  0.07, 0.02\}, \\
\vec{\epsilon}_b & =\{-0.06, 0.16, -0.22, 0.33, -0.26\}
\label{res_I:eqn}
\end{align}
contribute  2\% and 5\% respectively to the total  length of the $\gamma$ vectors.

@@

\color{orange}
The preferred values for $\kappa$ can be determined from the matrix elements of Eqn.~\ref{model11trans:eqn}: $\kappa^{-1}_1=2.82/1.15 =2.44$, $\kappa_2=-5.27/0.72=-7.31$.
The directions of the $\gamma$ vectors, and  the expectations from  \citetalias{1999PASP..111...63F} with $R=\kappa^{-1}$, are shown in 
Figure~\ref{plane:fig}.  There is almost perfect alignment between $\gamma^0$ and $R=2.44$ dust.
 From the Monte Carlo chains, we find a 68\% credible interval for $\kappa^{-1}_1$ of 2.26--2.63.

@@

\color{orange}
As a check of our analysis implementation, we generate ten simulated datasets with the same numbers and uncertainties as our data, based on a
SALT2 $x_1=0$, $c=0$ template and  \citetalias{1999PASP..111...63F} dust.  The supernovae
are assigned $E^F(B-V)$ parameters equal to the $B-V$ color excesses determined in our sample
(making an approximative correspondence between our observed top-hat filter colors and the 
model parameter $E^F(B-V)$) , and $R^F$ values drawn from a log-normal
distribution with a mean of $\ln{2.5}$ and 0.05 standard deviation.  The $\alpha$, $\beta$, and $\eta$ parameters are set to zero, so that dust and
measurement uncertainty are the only sources of variation in magnitudes.  We stack the results of the 10 simulations: the input $R^F=2.5$ is centered
within the 68\% credible interval for $\kappa^{-1}_1$ of 2.38--2.69.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/sn2014j.pdf} 
   \caption{Top: Measured color excess for SN~2014J and the best-fit predictions from this article.
   \color{orange}
   In addition to the prediction
   of our model, also shown is the prediction from the projection
   of our results onto the \citetalias{1999PASP..111...63F} extinction model.  Bottom: Residual color excess
   between the data and models, with a dotted line at zero shown for reference.
   \color{black}
   \label{sn2014j:fig}}
\end{figure}

The best-fit parameters for SN~2014J are 
%----
$e_0= 2.60$, $ e_1=-1.89$ with covariance
%----
\begin{equation}
\begin{pmatrix}
\begin{array}{rr}
0.050 & 0.058 \\
0.058 & 0.315
\end{array}
\end{pmatrix}.
\end{equation}
The predicted values of observed $E(B_o-V_o)$ from the fit are shown in Figure~\ref{sn2014j:fig}, where they are found to
overlap with the data within the error bars.

@@

\color{orange}
Our model can be associated with the extinction model of  \citetalias{1999PASP..111...63F} using Eqn.\ref{trans_I:eqn}
and the assumption that  SN~2011fe has zero extinction.  Transforming our parameters to those
of   \citetalias{1999PASP..111...63F}, we obtain
$A^F_V=0.77$ and $E^F(B-V)=1.28$ (which together imply $R^F=0.60$) with covariance
\begin{equation}
\begin{pmatrix}
\begin{array}{rr}
0.430 & -0.004 \\
-0.004 & 0.007
\end{array}
\end{pmatrix}.
\end{equation}
The color excesses predicted by these values are shown in Figure~\ref{sn2014j:fig}.

The imperfect correspondence between our model and that of   \citetalias{1999PASP..111...63F} is seen in the Figure.
While both models reproduce the observed color excesses at the extreme $U$ and $I$ bands, that of  \citetalias{1999PASP..111...63F} 
is unable to match the central colors.
The difficulty that  the  \citetalias{1999PASP..111...63F} model has in accounting for SN~2014J colors within the optical range is also evident
in Figure 3 of  \citet{2014ApJ...788L..21A}.  (They use
UV through NIR data to fit \citet{1999PASP..111...63F} dust with $R_V^F=1.4$ and $E(B-V)=1.37$.)
Our model predictions provide a significantly better match to the data
despite the only approximate correspondence of the synthetic SNfactory and \citet{2014ApJ...788L..21A} observer passbands.

@@

The resulting credible intervals of our parameters are given in Table~\ref{global2:tab}.  Comparison with the
results of the 2-parameter extrinsic model given in Table~\ref{global:tab} shows that  parameters that are common to both
are consistent well within 1$\sigma$ uncertainties.  The most significant change is in the slight reduction of the standard
deviations of the magnitude residuals $\sigma_X$.
The newly fitted $\gamma$ parameters have an almost identical association with the  \citetalias{1999PASP..111...63F} dust model, in that now
\begin{equation}
\begin{pmatrix}
 \frac{\gamma^0_X}{\gamma^0_1-\gamma^0_2} \\
\frac{\gamma^1_X}{\gamma^1_1-\gamma^1_2} 
\end{pmatrix}=
\begin{pmatrix}
\begin{array}{rr}
2.79 & 1.15 \\
-4.95 & 0.71
\end{array}
\end{pmatrix} 
\begin{pmatrix}
a(X) \\
b(X)
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_a \\
\vec{\epsilon}_b
\end{pmatrix},
\end{equation}
where the residual vectors 
\begin{align}
\vec{\epsilon}_a &=\{0.05, -0.06, -0.06,  0.06,  0.07\},\\
\vec{\epsilon}_b &=\{ -0.10, 0.20,  -0.19,  0.32, -0.30\}
\end{align}
\label{model11trans:eqn}
are perpendicular to the  \citetalias{1999PASP..111...63F} plane and
contribute similar 2\% and 4\% to the total  length of the vector.
(These can be compared to the results of Model I in Eqs.\ \ref{trans_I:eqn}--\ref{res_I:eqn}.)
%The effective total-to-selective extinction for the supernova sample has a slight shift to $R^F_V=2.23 \pm 0.16$.

@@

\color{black}
The analysis is validated using simulated datasets with and without third color parameters: the input parameters are recovered within  uncertainties.
We present here the results of one of the validation tests.  The input values of the supernova parameters are the median values from the analysis of real data.
Ten realizations of simulated data are generated based on the data covariances, and then analyzed separately.   The posteriors of the ten realizations
are aggregated and shown in Figures~\ref{simglobal1:fig}--\ref{simglobal5:fig}.  The input parameters are recovered with negligible bias
relative to the statistical uncertainties. 
\color{orange}
Note that some bias is expected: the priors for the distributions of $\Delta$, $k$, and $D$ are inconsistent
with the extracted distributions.  Also, 
there is a tendency for the slope parameters associated with the spectral features to be slightly underestimated, which is an expected consequence of 
the uncertainties in the predictor variable that lead to 
regression dilution \citep{spearman04}. 
\color{black}
